\chapter{Results and Analysis}
\label{cha:analysis}

\section{Introduction}
In this chapter we gather the results of the including the different
heuristics and algorithms in the solver. We go on to analyse these
results in order to determine the best combination of heuristic and
algorithm.

\section{Experimental Environment}
Here we define the set of levels on which we are going to compare the
heuristics and solvers. 

\subsection{Levelset}
The 3D engine (see Chapter \ref{cha:design}) we found came with a
number of simple levelsets. We chose the first of these as our primary
target set. This set of levels is consists mostly of small levels. The
average area of the levels in this was found to be $52.4$, whereas the
original levelset (of which the first floor-plan is depicted in Figure
\ref{fig:soko-org-screen}) has an average area of
$280.1$.\footnote{These areas is computed by multiplying the number of
  lines in the level-file with the length of the longest line in the
  level. So it is not exact, but is a good indicator of difficulty.}


\subsection{Methodology}
We will run all solvers on the level set defined above and observe the
number of total states discovered as a function of the solution trace.
These traces will give us some idea of which of these solvers, along
with which heuristics seem interesting.

We anticipate that the better (and thereby faster) the
solver/heuristic combination, the lower branching factor. The
branching factor for this kind of tree is would be calculated as an
average branching factor since it is irregular. We call this branching
factor, the \emph{determination} of the heuristic.

It is, however, important that the heuristic is not only determinate,
but also \emph{accurate}. A heuristic is accurate if it leads to
solution trace that is as short as possible.

\section{Speed vs. solution size}
\label{sec:speedvssolsize}
When we use a better heuristic the time needed to find a solution in a level is of course decreased. The solution found however is often bigger the faster it is found. Tables containing informations on the speed and solution size of the different heuristics can be found in appendix \ref{app:solutions}.

The heuristic names are simplified very much so that:
\begin{itemize}
\item +\{a, b, c\} is a HeuristicsAdder adding the heuristics a, b, and c
\item *\{a, b, c\} is a HeuristicsMultiplier multiplying the results of the heuristics a, b, and c
\item i is a SubGoalIndependence heuristic
\item c is a CornerHeursitic
\item 4 is a Box4x4Heuristic
\item B is a BoxOnGoalHeuristic
\item s is a ShortestPathHeuristic
\item a is a AveragePathHeuristic
\item A(a) is an \astar solver with the heuristic a
\item and finally a BFSolver is a BreadthFirst Solver
\end{itemize}

If we use the heuristic A(+\{*\{B,B,s,s\},c,4\})\footnote{an \astar solver with an HeuristicAdder (+) adding the results of a cornerHeuristic (c), a Box4x4Heuristic (4) and a HeuristicMultiplier (*) which multiplys the results of two BoxOnGoalHeursitics (B) and two ShortestPathHeuristic (s)} the time spent on finding a solution is much smaller than when we use the simple BreadthFirst Solver. And on the hard levels (Level40) only the best heuristics are able to find a solution in under 10 seconds.

At the same time the better heuristics make larger the solution. This is due to the fact that once the algorithm finds a path that looks good it follows it And the algorithm prefer to put a box all the way to a goal square before trying to get the next box to a goal.

The easier the level the less clear it is to see that the more advanced heursitics are faster. This is due to the increased time the advanced heuristics needs in order to evaluate each step. This increased time/step makes the evaluation time longer when the number of states seen are about the same for the advanced as for the the less advanced heuristic. On small, easy levels this increased step time matter alot, but as soon as the level increase in complexity the extra time spent is gained by not having to explore a lot of irrelevant states. This is clearly seen in Level40 where the A(+\{*\{s,s\},c,4\}) barely makes it in time and the best heuristic finds a solution in under a second. When the solution length gets even higher only the most advanced heuristics are able to find a solution. (See figure \ref{fig:std1solution}) As an example the solution we find to the 1st level of the original game (Figure \ref{fig:soko-org-screen} page \pageref{fig:soko-org-screen}) is on 380 steps which would be impossible for the BFSolver to find in a reasonable time.
\begin{figure}[htp]
\begin{verbatim}
		screen.1.level
		Name                Time      Sol. len. 
		----------------------------------------
		BFSolver                 10525         0
		A(+{*{i,i},c,4})         10001         0
		A(+{*{s,s},c,4})         10046         0
		A(+{*{B,B,s,s},c,4})       548       380
		A(+{*{B,B,a,a},c,4})       518       380
\end{verbatim}
	\caption{Result of running 5 different heuristics on the 1st level of the original levels}
	\label{fig:std1solution}
\end{figure}

\subsection{The long solution vs. the slow solution}
If we hold up the A(+\{*\{B,B,a,a\},c,4\}) against the BFSolver the level where the \astar solution is furthest away from the shortest solution (found by the BFSolver) is level04. The solution found by the \astar solver is 40 steps long while the shortest solution is only 23 steps long. That is an increase in solution size of 174\%.
The time used to find this solution is 28ms for the \astar and 1438ms for the BFSolver - that is only 2\% of the time the the BFSolver needed to find a solution.

This is the most extreme case found in the levels where the BFSolver could make it through in under 10seconds. Most of the time the solution found by the \astar is up to 125\% of the size of the shortest possible solution, and the time spent to find it is at most 50\% of the time spent finding the optimal solution.

\subsubsection{BFSolver only usable in small levels}
As the BFSolver is only able to find a solution in smaller levels: the biggest solution we have found using the BFSolver in 10second was on 107 steps. Had this been the size of the solution in the worst case scenario, namely a level which were an open field being 100+ squares from east to west and north to south, the amount of visited states would have reached $O({noPossibleDirections}^{size_{solution}} = 4^{107} = 2.6 \cdot 10^{64}$ visited states. This would be impossible for any computer to solve.

The \astar algorithm always tries to get closer to the goal. Its worst case running time is the same as the BFSolvers but for most of the time it will approach the goal state for every step it moves. At least if the heuristic assigned to the solver does a good estimation of the performance.

