\chapter{Design of Model and Solver}
\label{cha:design}

\section{Introduction}

\section{Overall model}
The planner can be devided into three areas: the objects describing
the world, the solvers (planning algorithms) and the heuristics used
by the \astar solver to estimate the performance of a given state.

\section{World objects}
Each level the planner is to solve is given in a level format
described on www.sokobano.de/wiki/. This level is loaded into an array
of Strings each letter being one square in the Sokobano world. These
strings are mainly used when the world is being explored by the
algorithm, but in order to do some more advanced moves the positions
of the boxes, the player and the goals is stored in a set giving
access to the positions.  When performing an action this is done by
the action itself. It is responsible for updating the world accroding
to what the effects of the action.

\section{The heuristics}
The heuristics are thoroughly described in chapter
\ref{cha:heuristics}.

\section{Solvers}
The planner has two solvers: a BreadthFirst (BF) Solver and an \astar
solver. At a given state the solvers look at all neighbour states,
states reachable using a single action, and adds them to a queue of
states to investigate.

\subsection{BreadthFirst Solver}
The BF solver adds the new found states to the end of the queue,
called unexplored states. This ensures that states discovered first
are explored before states found later. As a result it finds the
shortest possible solution, but it might take very long as it does not
care if states are close to or far from the goal state.

When a state is found is added to unexplored states only if it has not
been seen before. When it is first found it is added to the set
discovered states which is then able to tell if a state has been found
before.

\subsection{\astar solver}
The \astar solver works very much like the BF solver though instead of
a FIFO (first in, first out) queue it uses a priority queue sorting
the unexplored states according to their performance estimations
(determined by the heuristic). If a heuristic always returning the
same number the solver works as a BF solver because the only number
determining the priority of a state is the number of steps to get to
the state.

As some of the heuristics might throw a DeadlockException the \astar
solver also has to be able to handle this. If such an exception is
caught, the state is added to discovered states without being added to
the unexplored states. In this way the neighbour states to the
deadlocked state is not being explored making a solution building on a
deadlocked state (that we are able to find) impossible.

% Solver using A* (problem step 3)
